---
title: "Initial Korean covid 19"
author: "Marta Marriott"
date: "03/29/2024"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?

## Methods

Based on the problem description presents, the problem can be solved by using a categorical~categorical as we are comparing two categorical data types. The first is whether the patient is male or female, the second is whether the patient is dead or alive.

## Graphical results - use ggplot commands to do this  


```{r}
dd2 <- kRdata %>% group_by(result,sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
  
The two graphs show that women are contracting Covid at a higher rate. However, the mortality rate form the disease is highest among men.


## Numerical results

```{r}
table1 <- xtabs(~result + sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

The numerical data infers the conclusive results from the graphical data, the contraction rate of Covid is higher among women than men, but the mortality  percentage rate among men is  higher than women.

## Inferential results 




## Conclusions

